Exploring the Soil Microbiome coupling with the Crete island Data Cube

### 2024-04-04 manuscript.tex
add standards and integration in intro

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

@@

Crete is a continental forarc island \cite{ali2016}, fifth largest island of the mediterranean (~8300 km\textsuperscript{2}),
and a mediterranean biodiversity hotspot \cite{myers2000biodiversity}.
The island of Crete has been studied since the classical times for its'
fauna \cite{Sidiropoulos_Polymeni_Legakis_2017,Anastasiou2018Tenebrionid}, flora \cite{Krimbas_2005} and ecosystems \cite{Grove1993}.
Crete is home to the only endemic mammal of Greece, the Cretan shrew (\textit{Crocidura zimmermanni}),
more than 350 endemic arthropods \cite{bolanakis2024} and 183 endemic plants \cite{Kougioumoutzis2020}
among them a tree \textit{Zelkova abelicea}. Multifacet factors have shaped the
biodiversity of the island, for example the sharp elevation gradient \cite{trigas2013, FAZAN2017},
the complex evolutionary history \cite{POULAKAKIS2002} and the human - nature
interactions over thousands of years \cite{Vogiatzakis2008_med, Sfenthourakis2017}.
The major threats of human activities are becoming apperend in the island's ecosystems,
like desertification \cite{KARAMESOUTI2018266}, intesive grasing \cite{JouffroyBapicot2016},
climate change \cite{Kougioumoutzis2020,Vogiatzakis2016} and habitat loss \cite{ISPIKOUDIS1993259}.
Yet the topsoil microbial diversity of Crete has been unexplored.

@@

Data and metadata of these samplings are stored in different databases, yet 
great effort from these distinct communities have led to establishing standards
to enable FAIR data \cite{Wilkinson2016}. For amplicon sequences the Genome Standards
Consortium \cite{Field2011} has established the MIMARKS \cite{yilmaz2011minimum}
standards and has been a advocate for open and unrestricted data \cite{Amann2019}.
For soil data, rich metadata and platforms of hosting open data exist, e.g the 
European Soil Data Centre \cite{Panagos2022} and the World Soil Information
Service (WoSIS) \cite{Batjes2024}. These platforms also provide maps.

Spatial and remote sensing data

Soil structure and chemistry, holistic

@@

Interactomes have been characterised as a driver of community composition.
Microbial interactions challenges \cite{Faust2021} cohesion?

In this study, we integrade multiple types of data and methods to decipher hidden 
signals of the topsoil bacterial communities of Crete.

@@

Enumeratic all biodivarsity \cite{Anthony2023} and decipher the biogeochemical 
processes and interactions of fauna, flora and microbes \cite{Fry2019, Crowther2019,GRANDY201640,Delgado-Baquerizo2020}

Global hotspots \cite{Guerra2022}
and soil conservation instead of specific species \cite{Guerra2021}
and policy \cite{KONINGER2022} across countries \cite{Putten2023}

"A holistic perspective on soil architecture is needed as a key to soil functions"
\section{Conclusion}

>>>>

@@

Crete is a continental forarc island \cite{ali2016}, fifth largest island of the mediterranean (8350 km\textsuperscript{2}),
and a mediterranean biodiversity hotspot \cite{myers2000biodiversity}.
The island of Crete has been studied since the classical times for its'
fauna \cite{Sidiropoulos_Polymeni_Legakis_2017,Anastasiou2018Tenebrionid}, flora \cite{Krimbas_2005} and ecosystems \cite{Grove1993}.
Crete is home to the only endemic mammal of Greece, the Cretan shrew (\textit{Crocidura zimmermanni}),
more than 350 endemic arthropods \cite{bolanakis2024} and 183 endemic plants \cite{Kougioumoutzis2020}
among them a tree \textit{Zelkova abelicea}. Multifacet factors have shaped the
biodiversity of the island, for example the sharp elevation gradient \cite{trigas2013, FAZAN2017},
the complex evolutionary history \cite{POULAKAKIS2002} and the human - nature
interactions over thousands of years \cite{Vogiatzakis2008_med, Sfenthourakis2017}.
The major threats of human activities are becoming apperend in the island's ecosystems,
like desertification \cite{KARAMESOUTI2018266}, intesive grasing \cite{JouffroyBapicot2016},
climate change \cite{Kougioumoutzis2020,Vogiatzakis2016} and habitat loss \cite{ISPIKOUDIS1993259}.
Yet the topsoil microbial diversity of Crete has been unexplored.

@@

Data and metadata of these samplings are stored in different databases, yet 
great effort from these distinct communities have led to establishing standards
to enable FAIR data \cite{Wilkinson2016}. For amplicon sequences the Genome Standards
Consortium \cite{Field2011} has established the MIMARKS \cite{yilmaz2011minimum}
standards among others, and has been a advocate for open and unrestricted data \cite{Amann2019}.
Examples of rich metadata and platforms of hosting open data and digital soil maps are the 
European Soil Data Centre \cite{Panagos2022} and the World Soil Information
Service (WoSIS) of the ISRIC \cite{Batjes2024}. Apart from samplings,
spatial data are openly available terrestrial ecosystems.
Climatic data, land cover, desertification risk, aridity, soil type, normalized
vegetation index, bedrock geological formations. From the bacterial point of view, 
there curated databeses that classify in some baseline functionality.

Soil ecosystem functioning

Interactomes have been characterised as a driver of community composition soil
functioning \cite{GUSEVA2022108604}.
Still the challenges to infer actual microbial interactions remain \cite{Faust2021}.

@@

UN and EU soil goals for the 2030

In this study we ask: what the differences in the microbiome communities in different land use types?
Are there any climatic, geological, elevational, aridity or functional factors that affected the differences?
How the interactome changes over different land use types and climate?
Are there any distictions between arid regions of Crete?
To address these questions, we integrate multiple types of data and methods to decipher hidden 
signals of the Island Sampling Day soil microbiome data. In total, 144 samples from
72 sites (2 samples per site) of Crete are used with their metadata.
Warning signals are also identified for various types of ecosystems.

@@

Enumeratic all biodivarsity \cite{Anthony2023} and decipher the biogeochemical 
processes and interactions of fauna, flora and microbes \cite{Fry2019, Crowther2019,GRANDY201640,Delgado-Baquerizo2020}
Positive and negative in soils \cite{Liu2024}

Global hotspots \cite{Guerra2022}
and soil conservation instead of specific species \cite{Guerra2021}
and policy \cite{KONINGER2022} across countries \cite{Putten2023} and in Greece \cite{SCHISMENOS2022100035}

"A holistic perspective on soil architecture is needed as a key to soil functions" \cite{philippot2024the-interplay}
\section{Conclusion}
